Learning STRIPS action models with classical planning

### 2017-10-27 doc/icaps-18/strips-learning.tex
validation removed from evaluation of ICAPS18 paper

\begin{table}[hbt!]
	\begin{small}
	\begin{center}
		\begin{tabular}{l|l|l|l|l|l|l|}
			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
			\hline
			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7
		\end{tabular}
	\end{center}
	\end{small}
\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks.}
\label{tab:results_plans}
\end{table}

@@

\begin{table}[hbt!]
\begin{small}
	\begin{center}
		\begin{tabular}{l|c|c|c|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline
			Blocks & 0.04 & 0.00 & 72 \\
			Driverlog & 0.10 & 0.06 & 127 \\
			Ferry & 0.20 & 0.10 & 58 \\
			Floortile & 2.15 & 1.21 & 170 \\
			Gripper & 0.01 & 0.00 & 45 \\
			Miconic & 0.10 & 0.09 & 54 \\
			Satellite & 0.20 & 0.14 & 67 \\
			Transport & 0.20 & 0.18 & 60 \\
			Visitall & 0.43 & 0.39 & 37 \\
			Zenotravel & 2.78 & 2.76 & 70			
		\end{tabular}
	\end{center}
        \end{small}
	\caption{\small Planning time, preprocessing time and Plan length.}
	\label{tab:time_plans}	
\end{table}

@@

\begin{table}[hbt!]
\begin{small}
	\begin{center}
		\begin{tabular}{l|l|l|l|l|l|l|}
			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
			\hline 
		\end{tabular}
	\end{center}
	\end{small}
\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks exploiting the computed static predicates.}
\label{tab:results_plans_static}
\end{table}

\begin{table}[hbt!]
\begin{small}
	\begin{center}
		\begin{tabular}{l|c|c|c|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline 			
		\end{tabular}
	\end{center}
        \end{small}
	\caption{\small Planning time, preprocessing time and Plan length exploiting the computed static predicates.}
	\label{tab:time_plans_static}	
\end{table}

@@

\begin{table}[hbt!]
\begin{small}
	\begin{center}
		\begin{tabular}{l|l|l|l|l|l|l|}
			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
			\hline
			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Driverlog & 1.0 & 0.4 & 1.0 & 1.0 & 1.0 & 0.7 \\
			Ferry & 1.0 & 0.6 & 0.5 & 1.0 & 1.0 & 1.0 \\
			Floortile & 0.5 & 0.6 & 1.0 & 0.8 & 1.0 & 0.8 \\
			Gripper & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Miconic & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Satellite & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 1.0 \\
			Transport & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 0.5 \\
			Zenotravel & 1.0 & 0.5 & 0.6 & 0.6 & 1.0 & 1.0                         
		\end{tabular}
	\end{center}
	\end{small}
\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
\label{tab:results_plans_partial}
\end{table}

\begin{table}[hbt!]
\begin{small}
	\begin{center}
		\begin{tabular}{l|c|c|c|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline
			Blocks & 0.07 & 0.01 & 54  \\
			Driverlog & 0.06 & 0.03 & 71 \\
			Ferry & 0.18 & 0.14 & 50 \\
			Floortile & 0.81 & 0.68 & 92 \\
			Gripper & 0.01 & 0.00 & 37 \\
			Miconic & 0.06 & 0.04 & 40  \\
			Satellite & 0.18 & 0.12 & 43 \\
			Transport & 0.15 & 0.14 & 33 \\
			Zenotravel & 2.54 & 2.54 & 42                         
		\end{tabular}
	\end{center}
        \end{small}
	\caption{\small Planning time, preprocessing time and Plan length.}
	\label{tab:time_plans_partial}	
\end{table}

\subsection{Validating action models with classical planning}
Here we evaluate the performance of the compilation to validate whether a given action mode is compliant with a given set of observations. Table~\ref{tab:time_plans_val} reports the planning time, and prepossessing time used to solve the classical planning task that results from our compilation as well as the plan length of the found solutions.
\begin{table}[hbt!]
\begin{small}
	\begin{center}
		\begin{tabular}{l|c|c|r|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline
			Blocks & 0.00 & 0.00 \\
			Driverlog & 0.04 & 0.04 \\
			Ferry & 0.01 & 0.01 \\
			Floortile & 0.03 & 0.03 \\
			Gripper & 0.00 & 0.00 \\
			Miconic & 0.01 & 0.01 \\
			Satellite & 0.30 & 0.29 \\
			Transport & 0.02 & 0.02 \\
			Visitall & 0.07 & 0.07  \\
			Zenotravel & 11.65 & 11.39 			
		\end{tabular}
	\end{center}
        \end{small}
	\caption{\small Planning time, preprocessing time and Plan length for validating a given action model.}
	\label{tab:time_plans_val}	
\end{table}

@@

\subsubsection{Learning exploiting static predicates}
\subsubsection{Learning from partially specified action models}

@@

With regard to scalability of the compilation approach, the size of the compiled classical planning instances depends on the number of input examples. On the other hand the empirical results show that our approach, since is based on inferences instead of statistics, is able to generate non-trivial models from very small data sets. In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon. The same happens with all the actions for programming an action effect can also be applied in a single planning step.

>>>>

\begin{table}[hbt!]
	\begin{footnotesize}
	\begin{center}
		\begin{tabular}{l|l|l|l|l|l|l|}
			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
			\hline
			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7
		\end{tabular}
	\end{center}
	\end{footnotesize}
\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks.}
\label{tab:results_plans}
\end{table}

@@

\begin{table}[hbt!]
\begin{footnotesize}
	\begin{center}
		\begin{tabular}{l|c|c|c|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline
			Blocks & 0.04 & 0.00 & 72 \\
			Driverlog & 0.10 & 0.06 & 127 \\
			Ferry & 0.20 & 0.10 & 58 \\
			Floortile & 2.15 & 1.21 & 170 \\
			Gripper & 0.01 & 0.00 & 45 \\
			Miconic & 0.10 & 0.09 & 54 \\
			Satellite & 0.20 & 0.14 & 67 \\
			Transport & 0.20 & 0.18 & 60 \\
			Visitall & 0.43 & 0.39 & 37 \\
			Zenotravel & 2.78 & 2.76 & 70			
		\end{tabular}
	\end{center}
        \end{footnotesize}
	\caption{\small Planning time, preprocessing time and Plan length.}
	\label{tab:time_plans}	
\end{table}

@@

\begin{table}[hbt!]
\begin{footnotesize}
	\begin{center}
		\begin{tabular}{l|l|l|l|l|l|l|}
			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
			\hline
			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7                        
		\end{tabular}
	\end{center}
	\end{footnotesize}
\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks exploiting the computed static predicates.}
\label{tab:results_plans_static}
\end{table}

\begin{table}[hbt!]
\begin{footnotesize}
	\begin{center}
		\begin{tabular}{l|c|c|c|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline
			Blocks & 0.04 & 0.00 & 72 \\
			Driverlog & 0.10 & 0.06 & 127 \\
			Ferry & 0.20 & 0.10 & 58 \\
			Floortile & 2.15 & 1.21 & 170 \\
			Gripper & 0.01 & 0.00 & 45 \\
			Miconic & 0.10 & 0.09 & 54 \\
			Satellite & 0.20 & 0.14 & 67 \\
			Transport & 0.20 & 0.18 & 60 \\
			Visitall & 0.43 & 0.39 & 37 \\
			Zenotravel & 2.78 & 2.76 & 70                        
		\end{tabular}
	\end{center}
        \end{footnotesize}
	\caption{\small Planning time, preprocessing time and Plan length exploiting the computed static predicates.}
	\label{tab:time_plans_static}	
\end{table}

@@

\begin{table}[hbt!]
\begin{footnotesize}
	\begin{center}
		\begin{tabular}{l|l|l|l|l|l|l|}
			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
			\hline
			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Driverlog & 1.0 & 0.4 & 1.0 & 1.0 & 1.0 & 0.7 \\
			Ferry & 1.0 & 0.6 & 0.5 & 1.0 & 1.0 & 1.0 \\
			Floortile & 0.5 & 0.6 & 1.0 & 0.8 & 1.0 & 0.8 \\
			Gripper & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Miconic & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 1.0 \\
			Satellite & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 1.0 \\
			Transport & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 0.5 \\
			Zenotravel & 1.0 & 0.5 & 0.6 & 0.6 & 1.0 & 1.0                         
		\end{tabular}
	\end{center}
	\end{footnotesize}
\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
\label{tab:results_plans_partial}
\end{table}

\begin{table}[hbt!]
\begin{footnotesize}
	\begin{center}
		\begin{tabular}{l|c|c|c|}			 
			Domain & Total time & Preprocess & Plan length  \\
			\hline
			Blocks & 0.07 & 0.01 & 54  \\
			Driverlog & 0.06 & 0.03 & 71 \\
			Ferry & 0.18 & 0.14 & 50 \\
			Floortile & 0.81 & 0.68 & 92 \\
			Gripper & 0.01 & 0.00 & 37 \\
			Miconic & 0.06 & 0.04 & 40  \\
			Satellite & 0.18 & 0.12 & 43 \\
			Transport & 0.15 & 0.14 & 33 \\
			Zenotravel & 2.54 & 2.54 & 42                         
		\end{tabular}
	\end{center}
        \end{footnotesize}
	\caption{\small Planning time, preprocessing time and Plan length.}
	\label{tab:time_plans_partial}	
\end{table}

@@

@@

The size of the compiled classical planning instances depends on the number of input examples. The empirical results show that our approach can generate non-trivial models from very small data sets. In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon. The same happens with all the actions for programming an action effect can also be applied in a single planning step.
